---
output: github_document
---

# Statistical Rethinking, Homework, Week 1

## R packages

```{r, warning = FALSE, message = FALSE}
library(rethinking)
library(dplyr)
library(ggplot2)
```

## Exercise 1

* 4 water and 11 land observations
* Flat prior
* Grid approximation

```{r}
p_grid <- seq(0, 1, length.out = 1000)
prior_1 <- rep(1, 1000)
likelihood_1 <- dbinom(4, size = 4 + 11, prob = p_grid)
posterior_1_unstd <- likelihood_1 * prior_1
posterior_1 <- posterior_1_unstd / sum(posterior_1_unstd)

posterior_1_df <- tibble(
  p = p_grid,
  posterior_prob = posterior_1
)

ggplot(posterior_1_df, aes(p, posterior_prob)) +
  geom_line()
```

## Exercise 2

In the globe tossing example:

* 4 water and 2 land observations
* Prior that is zero below p = 0.5 and a constant above p = 0.5
* Grid approximation

```{r}
prior_2 <- ifelse(p_grid < 0.5, 0, 1)
likelihood_2 <- dbinom(4, size = 4 + 2, prob = p_grid)
posterior_2_unstd <- likelihood_2 * prior_2
posterior_2 <- posterior_2_unstd / sum(posterior_2_unstd)
posterior_2_df <- tibble(
  p = p_grid,
  posterior_prob = posterior_2
)

ggplot(posterior_2_df, aes(p, posterior_prob)) +
  geom_line()
```

## Exercise 3

For the posterior distribution from 2:

1. Compute 89% percentile and HPDI intervals.

```{r}
samples <- sample(p_grid, size = 1e4, replace = TRUE, prob = posterior_2)
(pi <- rethinking::PI(samples, prob = 0.89))
(hpdi <- rethinking::HPDI(samples, 0.89))
```

2. Compare the widths of these intervals. Which is wider? Why?

```{r}
(pi_width <- unname(pi[2] - pi[1]))
(hpdi_width <- unname(hpdi[2] - hpdi[1]))
## TODO Why is the PI wider?
```

3. If you only had the information in the interval, what might you misunderstand about the shape of the posterior distribution?

You may assume that the posterior distribution is symmetric and thus think that values below 0.5 are plausible when the model thinks they are not.

```{r}
df <- posterior_2_df %>%
  mutate(
    pi_interval = p > pi[1] & p < pi[2],
    hpdi_interval = p > hpdi[1] & p < hpdi[2]
  ) %>%
  arrange(desc(pi_interval))

df %>%
  arrange(p)

df %>%
  ggplot(aes(p, posterior_prob)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = 0,
      ymax = posterior_prob,
      fill = hpdi_interval
    )
  ) +
  scale_fill_manual(values = c("white", "firebrick")) +
  theme_minimal()
```

## Exercise 4

Assume that 1-in-5 (20%) of Water samples are accidentally recorded instead as "Land".

1. Write a generative simulation of this sampling process assuming that the true proportion of Water is 0.70.

```{r}
sample_unbiased <- sample(
  c("W", "L"),
  size = 1000,
  replace = TRUE,
  prob = c(0.7, 0.3)
)
table(sample_unbiased) / length(sample_unbiased)

sample_biased <- sample_unbiased
for (i in seq_along(sample_biased)) {
  if (sample_unbiased[i] == "W") {
    sample_biased[i] <- sample(
      c("W", "L"),
      size = 1,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  } 
}
table(sample_biased) / length(sample_biased)
```

```{r, eval = FALSE}
## TODO Why does this does not work?
sample_biased <- ifelse(
  sample_unbiased == "W",
  sample(c("W", "L"), size = 1, prob = c(0.8, 0.2)),
  sample_unbiased
)
```

2. Simulate 20 tosses with this sampling process:

```{r}
sample_unbiased <- sample(
  c("W", "L"),
  size = 20,
  replace = TRUE,
  prob = c(0.7, 0.3)
)
table(sample_unbiased) / length(sample_unbiased)

sample_biased <- sample_unbiased
for (i in seq_along(sample_biased)) {
  if (sample_unbiased[i] == "W") {
    sample_biased[i] <- sample(
      c("W", "L"),
      size = 1,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  } 
}
table(sample_biased)

W <- sum(sample_biased == "W")
p_grid <- seq(0, 1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(W, size = 20, prob = p_grid)
posterior_unstd <- likelihood * prior
posterior <- posterior_unstd / sum(posterior_unstd)

tibble(p = p_grid, posterior_prob = posterior) %>%
  ggplot(aes(p, posterior_prob)) +
  geom_line()

samples <- sample(p_grid, size = 1e4, replace = TRUE, prob = posterior)

```

